---
title: "p8105_hw1_msj2167"
author: "Michelle Jerry"
date: "2026-09-24"
output: github_document
---

Loading libraries 

```{r message = F}
library(tidyverse)
```

## Problem 1: 

### Data set description

To do this problem, I must load the tidyverse library and the penguins dataset

```{r}
data("penguins", package = "palmerpenguins")
```

The penguins data set has `r nrow(penguins)` rows and `r ncol(penguins)` columns. Each row is one penguin and the variables are `r names(penguins)`. 

Penguins were observed in one of three islands `r levels(penguins$island)` and belong to one of three species `r levels(penguins$species)`. They were seen between `r min(penguins$year)` and `r max(penguins$year)`. The physical measurements of penguins were recorded and included bill length and depth (mm), flipper length (mm), and body mass(g). The sex of the penguins were also recorded. 

The mean flipper length is
`r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)`

### Visualization of the dataset

Creating a scatterplot of bill length (x) and flipper length (y) by species 

```{r penguin_scatterplot, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) +
  geom_point()

ggsave("penguin_scatterplot.png")
```




